Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-30 papers/FX/draft.tex
Added pictures of scale estimates.

For convenience, we denote the scale parameter by $A$. It is the
constant $J_*(0)$ in the Taylor expansion \eqref{eq:scale_parameter}.
Hill \cite{hill1975simple} put forward an estimator of $A$ using his
tail index  estimator:
\[
\hat A = {k \over n} X_{(k+1)}^{\hat \alpha}
\]
where $X_{(k+1)}$ is the $(k+1)$-th upper order statistic of the
sequence $X_t$ as defined in section\ref{sec:Hill}.

@@

When applied to our problem, We would like to test, first of all,
whether there is a change of tail index in each of two 
series and, if there is no change in either series, whether the two
series have the same tail index. Obviously, for either question, if
the tail index changes, the extreme quantile will change.
For the first question, Hoga's test is directly applicable; for the
second question, we concatenate the two series before applying the
test.

@@

To have an idea about the power of Hoga's test, we run it on
concatenated samples drawn from t-distributions with different degrees
of freedom. Figure \ref{fig:t_sim_pair} shows the result.
\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth]{t_sim_pair.pdf}
  \caption{Test Statistics of concatenated t-samples with different
    degrees of freedom. Numbers on the axis are the degrees of freedom.}
  \label{fig:t_sim_pair}
\end{figure}
From the figure one can see that the power of the test decreases as
the minimum tail index of a concatenated pair increases.

@@

\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Energy_p.pdf}
                    \caption{``Energy''}
                    \label{fig:Energy_p}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Energy_alpha_K_ci.pdf}
                    \caption{``Energy''}
                    \label{fig:Energy_alpha_K_ci}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Consumer_Staples_p.pdf}
                    \caption{``Consumer Staples''}
                    \label{fig:Consumer_Staples_p}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Consumer_Staples_alpha_K_ci.pdf}
                    \caption{``Consumer Staples''}
                    \label{fig:Consumer_Staples_alpha_K_ci}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Information_Technology_p.pdf}
                    \caption{``Information Technology''}
                    \label{fig:Information_Technology_p}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Information_Technology_alpha_K_ci.pdf}
                    \caption{``Information Technology''}
                    \label{fig:Information_Technology_alpha_K_ci}
  \end{subfigure}
  \caption{\small Estimates of $p$ and $K$ of stocks in sectors of the
    S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals.
  }
\end{figure}
To compare the maximum likelihood estimates with the results obtained
from Hill's scale estimates, one can re-write the lower-tail
distribution function: For all $x < 0$,
\begin{eqnarray*}
  \P(X < x)
  &=&
  p {K^\alpha \over (K - x)^\alpha} \\
  &=&
  p {K^\alpha |x|^{-\alpha} \over (1 - K/x)^\alpha} \\
  &=&
  p K^\alpha |x|^{-\alpha} \left(
  1 + {\alpha K \over x} + O(|x|^{-2})
  \right)
\end{eqnarray*}
Comparing this expression with \eqref{eq:lower_tail_expansion}, one
can immediately identity $p K^\alpha$ as the scale parameter.

>>>>

For convenience, we denote the scale parameter by $A$. It is the
constant $J_*(0)$ in the Taylor expansion \eqref{eq:scale_parameter}.
Hill \cite{hill1975simple} put forward an estimator of $A$ using his
tail index  estimator:
\[
\hat A = {k \over n} X_{(k+1)}^{\hat \alpha}
\]
where $X_{(k+1)}$ is the $(k+1)$-th upper order statistic of the
sequence $X_t$ as defined in section\ref{sec:Hill}. $\hat \alpha$
is the Hill tail index estimator.
Estimates of $A$ of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in figure \ref{fig:Energy_scale},
\ref{fig:Consumer_Staples_scale} and
\ref{fig:Information_Technology_scale}, along with estimates obtained
using the hybrid method described in section
\ref{sec:hybrid_estimation}.

@@

When applied to our problem, We would like to test, first of all,
whether there is a change of tail index in each of two 
series and, if there is no change in either series, whether the two
series have the same tail index. Obviously, for either question, if
the tail index changes, the extreme quantile will change.
For the first question, Hoga's test is directly applicable; for the
second, we concatenate the two series before applying the
test.

@@

To have an idea about the power of Hoga's test, we run it on
concatenated samples drawn from t-distributions with different degrees
of freedom. Figure \ref{fig:t_sim_pair} shows the result.
\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth, trim={1cm 1cm 1cm 2cm},
  clip]{t_sim_pair.pdf}
  \caption{Test Statistics of concatenated t-samples with different
    degrees of freedom. Numbers on the axis are the degrees of freedom.}
  \label{fig:t_sim_pair}
\end{figure}
From the figure one can see that the power of the test decreases as
the minimum tail index of a concatenated pair increases.

@@

\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Energy_p.pdf}
                    \caption{``Energy''}
                    \label{fig:Energy_p}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Energy_alpha_K_ci.pdf}
                    \caption{``Energy''}
                    \label{fig:Energy_alpha_K_ci}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Consumer_Staples_p.pdf}
                    \caption{``Consumer Staples''}
                    \label{fig:Consumer_Staples_p}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Consumer_Staples_alpha_K_ci.pdf}
                    \caption{``Consumer Staples''}
                    \label{fig:Consumer_Staples_alpha_K_ci}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Information_Technology_p.pdf}
                    \caption{``Information Technology''}
                    \label{fig:Information_Technology_p}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Information_Technology_alpha_K_ci.pdf}
                    \caption{``Information Technology''}
                    \label{fig:Information_Technology_alpha_K_ci}
  \end{subfigure}
  \caption{\small Estimates of $p$ and $K$ of stocks in sectors of the
    S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals.
  }
\end{figure}
To compare the maximum likelihood estimates with the results obtained
from Hill's scale estimates, one can re-write the lower-tail
distribution function: For all $x < 0$,
\begin{eqnarray*}
  \P(X < x)
  &=&
  p {K^\alpha \over (K - x)^\alpha} \\
  &=&
  p {K^\alpha |x|^{-\alpha} \over (1 - K/x)^\alpha} \\
  &=&
  p K^\alpha |x|^{-\alpha} \left(
  1 + {\alpha K \over x} + O(|x|^{-2})
  \right)
\end{eqnarray*}
Comparing this expression with \eqref{eq:lower_tail_expansion}, one
can immediately identity $p K^\alpha$ as the scale parameter. Figure
\ref{fig:Energy_scale}, \ref{fig:Consumer_Staples_scale} and
\ref{fig:Information_Technology_scale} compare the estimates obtained
from Hill's scale estimator with those from the $p K^\alpha$ method.

\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
    \includegraphics[width=\textwidth]
                    {Energy_scale.pdf}
                    \caption{``Energy''}
                    \label{fig:Energy_scale}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
    \includegraphics[width=\textwidth]
                    {Consumer_Staples_scale.pdf}
                    \caption{``Consumer Staples''}
                    \label{fig:Consumer_Staples_scale}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
    \includegraphics[width=\textwidth]
                    {Information_Technology_scale.pdf}
                    \caption{``IT''}
                    \label{fig:Information_Technology_scale}
  \end{subfigure}
  \caption{\small Scale estimates obtained from Hill's scale estimator
    and from $p K^\alpha$ . Red dots are Hill's scale estimates and
    triangles are $p K^\alpha$ estimates.
  }
\end{figure}
